The Prevalence of E-Mail Header Injections on the World Wide Web

### 2016-05-20 Background/history.tex
Edits to background section.

\subsection{History of E-Mail Injection}

E-Mail Header Injection seems to have been first documented over a decade ago, in a late 2004 article on phpsecure.info~\cite{Tobozo} accredited to user \lstinline|tobozo@phpsecure.info| describing how this vulnerability existed in the reference implementation of the \texttt{mail()} function in PHP, and how it can be exploited. More recently, a blog post by Damon Kohler~\cite{DK} and an accompanying wiki article~\cite{Injection} describe the attack vector and outline a few defense measures for the same.

As this vulnerability was initially found in the \texttt{mail()} function of PHP, E-Mail Header Injection can be traced to as early as the beginning of the 2000's, present in the \texttt{mail()} implementation of PHP 4.0.

An example of the vulnerable code written in PHP is shown in Listing~\ref{code:phpemi}. This code takes in user input from the PHP superglobal \texttt{\$\_REQUEST[\textquotesingle email\textquotesingle]}, and stores it in the variable \texttt{\$from}, which is later passed to the \texttt{mail()} function to construct and send the e-mail.
\input{Background/php.tex}

When this code is given the malicious input \texttt{\lstinline{sai@sai.com\\nBCC:spc@spc.com}} as the value of the \texttt{\$\_REQUEST[\textquotesingle email\textquotesingle]}, it generates the SMTP Headers shown in Listing~\ref{code:smtpheaders}. It can be seen that the \texttt{CC} (carbon copy) header that we injected appears as part of the resulting SMTP message. This will make the e-mail get sent to the e-mail address specified as part of the \texttt{CC} as well.
\input{Background/smtpheaders}
%\input{Background/history_table}

>>>>

\subsection{History of \ehi}

We found the first \ehi description in a late 2004 article on phpsecure.info~\cite{Tobozo} accredited to user \lstinline|tobozo@phpsecure.info| describing how an \ehi vulnerability existed in the implementation of the \texttt{mail()} function in PHP and how it can be exploited. More recently, a blog post by Damon Kohler~\cite{DK} and an accompanying wiki article~\cite{Injection} describe the attack vector and outline few defense measures for \ehi vulnerabilities.

%% As this vulnerability was initially found in the \texttt{mail()} function of PHP, \ehi can be traced to as early as the beginning of the 2000's, present in the \texttt{mail()} implementation of PHP 4.0.

An example of the vulnerable code written in PHP is shown in Listing~\ref{code:phpemi}. This code takes in user input from the PHP superglobal \texttt{\$\_REQUEST[\textquotesingle email\textquotesingle]}, and stores it in the variable \texttt{\$from}, which is later passed to the \texttt{mail()} function to construct and send the e-mail.

\input{Background/php.tex}

% TODO: Change all domains to example.com
% TODO: Sai, this is the first time we mention SMTP, need more background and a reference to the RFC.

\textbf{Adam: This example is not correct, the STMP headers in listing 2 should have the "From" headers from the injection.}

When this code is given the malicious input \texttt{\lstinline{sai@sai.com\\nCC:spc@spc.com}} as the value of the \texttt{\$\_REQUEST[\textquotesingle email\textquotesingle]}, it generates the SMTP Headers shown in Listing~\ref{code:smtpheaders}. It can be seen that the \texttt{CC} (carbon copy) header that we injected appears as part of the resulting SMTP message. This will make the e-mail get sent to the e-mail address specified as part of the \texttt{CC} as well. 
\input{Background/smtpheaders}
%\input{Background/history_table}

### 2016-05-20 Background/impact.tex
Edits to background section.

\subsection{Potential Impact}

The impact of the vulnerability can be pretty far-reaching.
Table~\ref{tab:usage} shows the current server-side language usage statistics on the Web~\cite{W3techs}.

\begin{table}[!tb]
	\centering
	\begin{tabular}{|p{4cm}|p{4cm}|}
		\hline
		\multicolumn{1}{|c|}{\textbf{Server Side Language}} & \multicolumn{1}{c|}{\textbf{\% of Usage}}\\
		\hline
		PHP & 81.9\\
		\hline
		Java & 3.1\\
		\hline
		Ruby & 0.6\\
		\hline
		Python & 0.2\\
		\hline

	\end{tabular}
	\caption[\titlecap{Language usage statistics}]{Language usage statistics compiled from w3techs~\cite{W3techs}.}
	\label{tab:usage}
\end{table}

PHP, Java, Python, and Ruby (combined) account for over 85\%\,\footnotemark{} of the websites measured. The vulnerability can be exploited to do potentially any of the following:

\footnotetext{A website may use more than one server-side programming language}

\begin{itemize}
	\item Phishing and Spoofing Attacks\\
    Phishing~\cite{phishing} (a variation of spoofing~\cite{spoofing_attack}) refers to an attack where the recipient of an e-mail is made to believe that the e-mail is a legitimate one. The e-mail usually redirects them to a malicious website, which then steals their credentials.

    E-Mail Header Injection gives attackers the ability to inject arbitrary headers into an e-mail sent by a website and control the output of the e-mail. This adds credibility to the generated e-mail, as it is sent right from the websites and people are more ready to trust e-mail that is received from the website directly and can thus result in more successful phishing attacks.

	\item Spam Networks\\
	Spam networks can use E-Mail Header Injection vulnerabilities on the ability to send a large amount of e-mail from servers that are trusted. By adding additional \texttt{cc} or \texttt{bcc} headers to the generated e-mail, attackers can easily achieve this effect.

	Due to the e-mails being from trusted domains, recipient e-mail clients might not flag them as spam. If they do flag them as spam, then that can lead to the website being blacklisted as a spam generator.

	\item Information Extraction of legitimate users\\
	E-Mails can contains sensitive data that is meant to be accessed only by the user. Due to E-Mail Header Injection, an attacker can easily add a \texttt{bcc} header, and send the e-mail to himself, thereby extracting important information.
	User privacy can thus be compromised, and loss of private information can by itself lead to other attacks.

	\item Denial of service by attacking the underlying mail server\\
    Denial of service attacks (DoS), can also be aided by E-Mail Header Injection. The ability to send hundreds of thousands of e-mails by just injecting one header field can result in overloading the mail server, and cause crashes and/or instability. 
\end{itemize}

It is evident that E-Mail Header Injection is a critical vulnerability that web applications must address.

>>>>

\subsection{Impact of \ehi}

The impact of an \ehi vulnerability can be far-reaching. According to
w3tech, PHP, Java, Python, and Ruby (combined) account for over
85\%\,\footnotemark{} of the server-side programming languages in
websites measured, and the default implementation of the \email functionality of these languages is vulnerable to \ehi.

\footnotetext{A website may use more than one server-side programming language.}

%% \begin{table}[!tb]
%% 	\centering
%% 	\begin{tabular}{|p{4cm}|p{4cm}|}
%% 		\hline
%% 		\multicolumn{1}{|c|}{\textbf{Server Side Language}} & \multicolumn{1}{c|}{\textbf{\% of Usage}}\\
%% 		\hline
%% 		PHP & 81.9\\
%% 		\hline
%% 		Java & 3.1\\
%% 		\hline
%% 		Ruby & 0.6\\
%% 		\hline
%% 		Python & 0.2\\
%% 		\hline

%% 	\end{tabular}
%% 	\caption[\titlecap{Language usage statistics}]{Language usage statistics compiled from w3techs~\cite{W3techs}.}
%% 	\label{tab:usage}
%% \end{table}

An \ehi vulnerability can be exploited to do potentially any of the
following:

\noindent \textbf{Phishing and Spoofing Attacks} 
    Phishing~\cite{phishing} (a variation of spoofing~\cite{spoofing_attack}) refers to an attack where the recipient of an \email is made to believe that the \email is  legitimate when it was really created by a malicious party. The \email usually redirects the victim to a malicious website, which then steals their credentials or infects their computer with malware (via a drive-by-download).

    \ehi gives attackers the ability to inject arbitrary headers into an \email sent by a website \emph{and control the output of the \email}. This adds credibility to the generated \email, as it is sent from the website's mail server and users (and anti-spam defenses) are more likely to trust an e-mail that is received from the proper mail server. Therefore, attackers could leverage \ehi vulnerabilities to perform enhanced phishing attacks.

\noindent\textbf{Spam Networks}
	Spam networks can use \ehi vulnerabilities to send a large amount of \email from servers that are trusted. By adding additional \texttt{cc} or \texttt{bcc} headers to the generated e-mail, attackers can easily choose the recipient of the spam email.

	Due to the \email being from trusted domains, recipient \email clients and anti-spam systems might not flag them as spam. If they do flag them as spam, then that can lead to the website being blacklisted as a spam generator (which would cause a Denial of Service on the vulnerable web application).

\noindent\textbf{Information Extraction}
	\emails can contain sensitive data that is meant to be accessed only by the user. Due to an \ehi vulnerability, an attacker can add a \texttt{bcc} header, and the \email server will send a copy of the private \email to the attacker, thereby extracting important information.
	User privacy can thus be compromised, and loss of private information can by itself lead to other attacks.

    \noindent\textbf{Denial of Service}
    Denial of service attacks (DoS), can also be caused by exploiting an \ehi vulnerability. The ability to send many \emails by injecting one header field can result in overloading the mail server and cause crashes or instability.

%It is evident that E-Mail Header Injection is a critical vulnerability that web applications must address.

### 2016-05-20 Background/lang_aff.tex
Edits to background section.

\subsection{Languages Affected}
\label{languages}
This section describes the popular languages which exhibit this type of vulnerability. This section is not intended as a complete reference of vulnerable functions and methods, but rather as a guide that specifies which parts of the language are known to have the vulnerability.

\begin{itemize}
\item{\textbf{PHP}} - PHP was one of the first languages found to have this vulnerability in its implementation of the \texttt{mail()} function. The early finding of this vulnerability can be attributed in part to the success and popularity of the language for creating web pages. According to w3techs \cite{W3techs}, PHP is used by 81.9\% of all the websites in existence, thereby creating the possibility of this vulnerability to be widespread.

After 13 further iterations of the language since the 4.0 release (the current version is 7.1), the \texttt{mail()} function is yet to be fixed after 15 years. However, it is specified in the PHP documentation \cite{PHPDocs} that the \texttt{mail()} function does not protect against this vulnerability.
A working code sample of the vulnerability, written in PHP~5.6 (latest well-supported version), is shown in  Listing~\ref{code:phpemi}.

\item{\textbf{Python}} -
A bug was filed about the vulnerability in Python's implementation of the \texttt{email.header} library and its header parsing functions allowing newlines in early 2009, which was followed up with a partial patch in early 2011.

Unfortunately, the bug fix was only for the \texttt{email.header} package, and thus is still prevalent in other frequently used packages such as \texttt{email.parser}, where both the classic \texttt{Parser()} and the newer \texttt{FeedParser()} exhibit the vulnerability even in the latest versions - \texttt{2.7.11} and \texttt{3.5}. The bug fix was also not backported to older versions of Python.
There is no mention of the vulnerability in the Python documentation for either library.
A working code sample of the vulnerability, written in Python 2.7.11, is shown in Listing~\ref{code:pyemi}.
\input{Background/py.tex}

\item{\textbf{Java}} -
Java has a bug report about E-Mail Header Injection filed against its JavaMail API. A detailed write-up by Alexandre Herzog \cite{Herzog.2014} is complete with a proof of concept program that exploits the API to inject headers.

\item{\textbf{Ruby}} -
From our preliminary testing, Ruby's built-in Net::SMTP library has this vulnerability. This is not documented on the library's homepage.
%A working code sample of the vulnerability, written in Ruby 2.0.0 (the latest stable version at the time of writing), is shown in Listing~\ref{code:rubyemi}.

@@

\end{itemize}

>>>>

\subsection{Languages Affected}
\label{languages}
This section describes the popular programming languages that contain \ehi vulnerabilities in their standard email libraries. This section is not intended as a complete reference of vulnerable functions and methods, but rather as a guide that specifies which parts of the language are vulnerable.

\noindent{\textbf{PHP}} was one of the first languages found to be vulnerable to \ehi in its implementation of the \texttt{mail()} function at the time of release of PHP~4.0. The early finding of this vulnerability can be attributed in part to the success and popularity of the language for creating web pages. According to w3techs~\cite{W3techs}, PHP is used by 81.9\% of all the websites.

After 13 further iterations of PHP since the 4.0 release (the current version is 7.1), the \texttt{mail()} function is yet to be fixed after 15 years. However, it is specified in the PHP documentation~\cite{PHPDocs} that the \texttt{mail()} function does not protect against \ehi.
A working code sample of the vulnerability, written in PHP~5.6 (latest well-supported version), is shown in  Listing~\ref{code:phpemi}.

\noindent{\textbf{Python}}
A bug was filed about an \ehi vulnerability in Python's implementation of the \texttt{email.header} library and the header parsing functions allowing newlines in early 2009, which was followed by a partial patch in early 2011.

Unfortunately, the bug fix was only for the \texttt{email.header} package, and thus exists in other frequently used packages such as \texttt{email.parser}, where both the classic \texttt{Parser()} and the newer \texttt{FeedParser()} contain \ehi vulnerabilities even in the latest versions: \texttt{2.7.11} and \texttt{3.5}. The bug fix was also not backported to older versions of Python.
There is no mention of the vulnerability in the Python documentation for either library.
A working code sample of the vulnerability, written in Python 2.7.11, is shown in Listing~\ref{code:pyemi}.

\input{Background/py.tex}

\noindent{\textbf{Java}} has a bug report about \ehi filed against its \texttt{JavaMail} API. A detailed write-up by Alexandre Herzog~\cite{Herzog.2014} contains a proof-of-concept program that exploits the API to inject headers.

\noindent{\textbf{Ruby}} 
From our preliminary testing Ruby's built-in \texttt{Net::SMTP} library also has an \ehi library. This is not documented on the library's homepage.
%A working code sample of the vulnerability, written in Ruby 2.0.0 (the latest stable version at the time of writing), is shown in Listing~\ref{code:rubyemi}.

@@

### 2016-05-20 Background/pbm_bg.tex
Edits to background section.

E-Mail Header Injection belongs to a broad class of vulnerabilities known simply as injection attacks. However, unlike its more popular siblings, SQL injection~\cite{sql1}, \cite{sql0}, \cite{sql2}, Cross-Site Scripting (XSS)~\cite{Injection1}, \cite{KleinAmit} or even HTTP Header Injection~\cite{sessionride}, relatively little research is available on E-Mail Header Injection.

As with other vulnerabilities in this class, E-Mail Header Injection is caused due to improper sanitization (or lack thereof) of user input. If the script that constructs e-mails from user input fails to check for the presence of e-mail headers in the user input, a malicious user --- using a well-crafted payload --- can control the headers set for this particular e-mail. This can be leveraged to enable malicious attacks, including, but not limited to, spoofing, phishing, etc.

>>>>

\ehi belongs to a broad class of vulnerabilities known as command injection vulnerabilities. However, unlike its more popular siblings, SQL injection~\cite{sql1, sql0, sql2}, Cross-Site Scripting~\cite{Injection1, KleinAmit}, or HTTP Header Injection~\cite{sessionride}, relatively little research is available on \ehi vulnerabilities.

As with other vulnerabilities in this class, \ehi is caused due to improper or nonexistent sanitization of user input. If the program that constructs \emails from user input fails to check for the presence of \email headers in the user input, a malicious user---using a well-crafted payload---can control the headers set for this particular e-mail. \ehi vulnerabilities can be leveraged to enable malicious attacks, including, but not limited to, spoofing or phishing.

### 2016-05-20 Background/php.tex
Edits to background section.

\lstset{language=PHP,caption={PHP program with e-mail header injection vulnerability.},label={code:phpemi}}
\begin{lstlisting}
$from = $_REQUEST['email'];
$subject = 'Hello XYZ';
$message = 'We need you to reset your password';
$to = 'xyz@xyz.edu';

>>>>

\begin{lstlisting}[language=PHP,caption={PHP program with e-mail
      header injection vulnerability.},label={code:phpemi}, float]
$from = $_REQUEST['email'];
$subject = 'Hello XYZ';
$message = 'We need you to reset your password';
$to = 'xyz@xyz.edu';

### 2016-05-20 Background/py.tex
Edits to background section.

\lstset{language=Python,caption={Python program with e-mail header injection vulnerability.},label={code:pyemi}}
\begin{lstlisting}
from email.parser import Parser
import cgi
form = cgi.FieldStorage()
to = form["email"] # input() exhibits 
# the same behavior
msg = """To: """ + to + """\n
From: <user@example.com>\n
Subject: Test message\n\n
Body would go here\n"""

>>>>

\begin{lstlisting}[language=Python,caption={Python program with e-mail
      header injection vulnerability.},label={code:pyemi}, float]
from email.parser import Parser
import cgi
form = cgi.FieldStorage()
to = form["email"] # input() exhibits 
# the same behavior
msg = """To: """ + to + """\n
From: <user@example.com>\n
Subject: Test message\n\n
Body would go here\n"""

### 2016-05-20 Background/ruby.tex
Edits to background section.

\lstset{language=Ruby,caption={Ruby program with e-mail header injection vulnerability.},label={code:rubyemi}}
\begin{lstlisting}
require 'sinatra'
require 'net/smtp'

>>>>

\begin{lstlisting}[language=Ruby,caption={Ruby program with e-mail
      header injection vulnerability.},label={code:rubyemi}, float]
require 'sinatra'
require 'net/smtp'

### 2016-05-20 Discussion/limits.tex
Edits to background section.

       	\lstset{language=HTML,caption={E-Mail field checker - false positives, the system incorrectly classifies this as an e-mail form.},label={code:false_positive}}
       	\begin{lstlisting}
<form method="post">
E-Mail us if you have any questions!!
<input type="text" name="query"><br>
<input type="submit" value="Search">
</form>
       	\end{lstlisting}

@@

\lstset{language=HTML,caption={HTML page with e-mail form, written in a different language - French.},label={code:htmlfrench}, literate=%
	{é}{{\'e}}1}
\begin{lstlisting}
<!doctype html>
<html lang="fr">
<head>
<meta charset="utf-8">
<meta name="author" content="XYZ">
<title>Mock Email</title></head>
<body>
<form action="{path-to-back-end}" method="post">
<input type="text" placeholder="courrier électronique"
	name="courrier_électronique"><br>
<textarea name="msg" rows="20" cols="120"></textarea>
<input type="submit" value="courrier électronique!">
</form></body>
</html>
\end{lstlisting}

>>>>

       	\begin{lstlisting}[language=HTML,caption={E-Mail field checker
              - false positives, the system incorrectly classifies
              this as an e-mail form.},label={code:false_positive}, float]
<form method="post">
E-Mail us if you have any questions!!
<input type="text" name="query"><br>
<input type="submit" value="Search">
</form>
       	\end{lstlisting}

@@

\begin{lstlisting}[language=HTML,caption={HTML page with e-mail form, written in a different language - French.},label={code:htmlfrench}, literate=%
	{é}{{\'e}}1, float]
<!doctype html>
<html lang="fr">
<head>
<meta charset="utf-8">
<meta name="author" content="XYZ">
<title>Mock Email</title></head>
<body>
<form action="{path-to-back-end}" method="post">
<input type="text" placeholder="courrier électronique"
	name="courrier_électronique"><br>
<textarea name="msg" rows="20" cols="120"></textarea>
<input type="submit" value="courrier électronique!">
</form></body>
</html>
\end{lstlisting}

### 2016-05-20 System/proof_of_concept.tex
Edits to background section.

\lstset{language=HTML,caption={HTML page for showcasing e-mail header injection, a simple front-end for our examples.},label={code:html}}
\begin{lstlisting}
<!doctype html>
<html lang="en">
<head>
<meta charset="utf-8">
<meta name="author" content="XYZ">
<title>Mock Email</title>
</head>
<body>
<form action="{path-to-back-end}" method="post">
<input type="text" placeholder="Email" name="email" id="e-mail"><br>
<textarea name="msg" rows="20" cols="120"></textarea>
<input type="submit" value="Email Me!">
</form>
</body>
</html>
\end{lstlisting}

>>>>

\begin{lstlisting}[language=HTML,caption={HTML page for showcasing
      e-mail header injection, a simple front-end for our
      examples.},label={code:html}, float]
<!doctype html>
<html lang="en">
<head>
<meta charset="utf-8">
<meta name="author" content="XYZ">
<title>Mock Email</title>
</head>
<body>
<form action="{path-to-back-end}" method="post">
<input type="text" placeholder="Email" name="email" id="e-mail"><br>
<textarea name="msg" rows="20" cols="120"></textarea>
<input type="submit" value="Email Me!">
</form>
</body>
</html>
\end{lstlisting}
